### 2009-08-12 doc/tech_doc/sec/arc-security-documentation.tex
more change for security documentation

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {Weizhong Qiang},
  pdftitle = {Documentation of the ARC security framework},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Technical Description} % (fold)
\label{cha:tech_description}

\section{Introduction} % (fold)
\label{sec:introduction}

@@

In the implementation of the ARC1, there is a Service Container – the Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting place for various services in application level, as well as a flexible and efficient communication mechanism.
HED contains a framework for implementing and enforcing authentication and authorization. Each Message Chain Component (MCC) or service has a common interface for implementing various authentication and authorization functionality. This functionality is implemented by using pluggable components (plug-ins) called SecHandler. The SecHandler components are C++ classes and provide method for processing messages traveling through Message Chains of the HED. Each MCC or Service  usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called ``incoming'' and ``outgoing'' respectively. It is possible for MCC or Service to implement other set of queues. Please check documentation of particular component for that particular information. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well.
Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs. Some of implemented SecHandler components also make use of pluggable and configurable sub-modules which specifically handle various security functionalities, such as authorization, authentication, etc. The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents. Figure 1 gives the structure of a MCC/Service, and the message sequence inside it. And Figure 2 shows the configuration of SecHandler components for an example ``Echo'' service.

@@

When one component (MCC or service) is loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well. 
There is one simple interface (see Figure 3) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.

@@

Figure 4 shows the definition of abstract class PDP. The implementation could be some function which implements the interface by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result, or just by composing the policy evaluation request, invoking some remote policy decision web service and getting back the evaluation result.

@@

@@

    \item pdpservice.invoker – PDP Service Invoker
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation engine can be accessed remotely and maintained centrally.

@@

Figure 5 shows the UML class diagram about the policy evaluation engine. It shows all classes and relations simultaneously for getting the overall picture. 
The Evaluator class is the key class for policy evaluation. It accepts request evaluates it against loaded policy and returns evaluation response.

Three abstract factories - FnFactory, AlgFactory, AttributeFactory - are responsible for creating the Function, CombiningAlg and AttributeValue objects correspondingly. The classes inherited from CombiningAlg class take care of implementing various combining algorithms which define relations between \textit{$<$Rule/$>$} elements in policy. The AttributeValue type of classes are used for processing different types of \textit{$<$Attribute/$>$} and similar elements. The Function classes take care of comparing \textit{$<$Attribute/$>$} elements of request and policy.

The Policy class parses \textit{$<$Policy/$>$} or \textit{$<$Rule/$>$} elements and creates  CombingAlg objects according to the \textit{$<$RuleCombiningAlg/$>$} attribute of \textit{$<$Policy/$>$}, Function objects according to the \textit{$<$Function/$>$} attribute of \textit{$<$Attribute/$>$} and AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. Those objects will be used when evaluating the request.

The Request class is responsible for parsing \textit{$<$Request/$>$} element and creates corresponding AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. When evaluating, each AttributeValue in request will be evaluated against corresponding AttributeValue in the policy by using relevant Function.

Due to extensible architecture of code it is relatively easy to add support for new types of  AttributeValue, Function and CombingAlg objects in this way supporting various types of XML based policy languages.

\subsection{Schemas for policy evaluation engine} % (fold)
\label{subsec:policy_schema}
The schema for ARC Policy is available at

@@

The hierarchy tree of ARC Policy is show below (numbers show multiplicity of elements)

@@

\subsection{Basic Elements of Policy} % (fold)
\label{subsec:poicy_element}
There are 2 basic objects - "policy" and "request". There is 1 main actor - Evaluator. Curretly there are two types of elements in policy: Policy and Rule. Policy element is made of Rule elements.
Evaluator matches request to policy and produces one of 4 following results:

@@

\subsection{Policy Matching} % (fold)
\label{subsec:poicy_matching}
Policy is made of Rule elements. Request is evaluated against each Rule. Each evaluation produces same results as policy evaluation described above. The results from all Rules are then combined in order to produce final result for whole policy. Results Combining Algorithm is specified in Policy. There are 26 algorithms currently:

@@

\subsection{Request Structure} % (fold)
\label{subsec:request_structure}

@@

\subsection{Rule Composition and Matching} % (fold)
\label{subsec:rule_comp_match}

@@

\subsection{Rule Elements Matching} % (fold)
\label{subsec:rule_element_match}
For elements without attributes those elements have:

@@

@@

For making usage of policy evaluation engine more convenient basic Evaluator class is complemented by additional interfaces. Below are examples of steps needed to carry out policy evaluation and corresponding helper interfaces.

@@

\begin{verbatim}
    // Create object which provides an interface 
    // for loading other objects
    ArcSec::EvaluatorLoader eval_loader; 
    //Load the Evaluator 
    ArcSec::Evaluator* eval = NULL;
    // Define name of policy evaluator.
    // This one is for evaluation ARC policies
    std::string evaluator = "arc.evaluator"; 
    eval = eval_loader.getEvaluator(evaluator);
\end{verbatim}
b)Create the policy object:

\begin{verbatim}
    ArcSec::Policy* policy = NULL; 
    // Define type of policy – ARC policy in this case
    std::string policyclassname = "arc.policy"; 
    // Define source from which policy to be taken
    ArcSec::SourceFile policy_source("Policy_Example.xml"); 
    // Load and parse policy
    policy = eval_loader.getPolicy(policyclassname, policy_source); 
\end{verbatim}

c)Create the request:
\begin{verbatim}
    ArcSec::Request* request = NULL;
    // Define type of request – ARC request in this case
    std::string requestclassname = "arc.request";
    // Define source from which request to be taken
    ArcSec::SourceFile request_source("Request.xml");
    // Load and parse request
    request = eval_loader.getRequest(requestclassname, request_source);
\end{verbatim}

@@

The Evalutor::evaluate() method can also be feed up with both Policy/Request objects and their sources in any combination. See example code at http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/testinterface.cpp for more details about usage of the interface.

@@

\section{Policy Decision Service} % (fold)
\label{sec:policy_decision_service}
Policy decision service is a service implementation which contains the functionality of ArcPDP. It will accept the soap request containing policy decision request and return soap response containing policy decision response.

@@

 http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl.

@@

\section{Security Attributes} % (fold)
\label{sec:sec_attributes}

@@

Security Attributes represent security related information inside HED framework and store information representing various aspects needed to perform authorization decison - identity of client, requested action, targeted resource, constraint policies. 
Each kind of Security Attribute is represented by own class inherited from parent SecAttr class <arc/message/SecAttr.h>. Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using Export() method. Currently only supported format is ARC Policy/Request XML document described in Section 8.1and 8.2
Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth <arc/message/MessageAuth.h> and MessageAuthContext <arc/message/Message.h> storing collected attributes per request and per session correspondingly. Each attribute is assigned a name. Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins. See Section 5.2for available Collectors and corresponding Security Attributes. 
Note for service developers: Services may implement own authorization algorithms. But they may use Security Atributes as well by providing instances of classes inherited from SecAttr and running them through either configured or hardcoded processors/PDPs.
Processors of Security Attributes are implemented as Policy Decision Point components. Currently there are 2 PDP components available:
Arc PDP makes use of Security Attributes containing identities of client, resource and requested action. It evaluates either all or selected set of attributes against specified Policy documents thus making it possible to enforce policies defined/selected by service providers. 
Delegation PDP is described below in Section 6.3.

@@

\subsection{Delegation Interface} % (fold)
\label{subsec:delegation_interface}
Delegation interface in the ARC1 is implemented using Web Service approach. Each ARC1 service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials. Here is how delegation procedure works (also shown in figures 6 and 7) :

@@

@@

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath> 
            <tls:CACertificatePath>./cacert.pem</tls:CACertificatePath> 
              <tls:CACertificatesDir>./certificates</tls:CACertificatesDir> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="soap">POST</next> 
        </Component> 
        <Component name="soap.service" id="soap"> 
            <next id="echo"/> 
        </Component> 
        <Service name="echo" id="echo"> 
            <SecHandler name='delegation.handler' id='delegation' event='incoming'> 
              <deleg:Type>x509</deleg:Type> 
              <deleg:Role>delegatee</deleg:Role> 
              <deleg:DelegationServiceEndpoint>https://glueball.uio.no:60000/delegation</deleg:DelegationServiceEndpoint> 
              <!--The credential and trusted certificates which will be used by this security handler to contact with delegation service -->
              <deleg:KeyPath>./testkey-nopass.pem</deleg:KeyPath> 
              <deleg:CertificatePath>./testcert.pem</deleg:CertificatePath> 
              <deleg:CACertificatesDir>./certificates</deleg:CACertificatesDir> 
              <deleg:CACertificatePath>./testcacert.pem</deleg:CACertificatePath> 
            </SecHandler> 
            <next id="echo"/> 
            <echo:prefix>[ </echo:prefix> 
            <echo:suffix> ]</echo:suffix> 
        </Service> 
    </Chain> 
\end{verbatim}

@@

\begin{verbatim}
     Arc::XMLNode sechanlder_nd("\ 
        <SecHandler name='delegation.handler' id='delegation' event='outgoing'>\ 
          <Type>x509</Type>\ 
          <Role>delegator</Role>\ 
          <!--The endpoint of delegation service, to which this security handler will delegate the credential-->\
          <DelegationServiceEndpoint>https://glueball.uio.no:60005/delegation</DelegationServiceEndpoint>\ 
          <!--The endpoint of the peer service, to which this WS client will access-->\
          <PeerServiceEndpoint>https://selectron.uio.no:60000/echo</PeerServiceEndpoint>\ 
         <!--The key and certificate of the credential, which is be delegated to the delegation service-->\
          <KeyPath>./testkey-nopass.pem</KeyPath>\ 
          <CertificatePath>./testcert.pem</CertificatePath>\ 
         <!--Another option of defining the location of credential-->\
          <!--ProxyPath>/tmp/5612d050.pem</ProxyPath-->\ 
         <!--Yet another option of defining how to get the credential; this option is useful for the service which has already acquired the delegated credential (this delegated credential is stored under /tmp/hash_value.pem) from the delegation service, yet still needs to delegate this credential to another delegation service; the DelegationCredentialIdentity can be parsed from the tls connection from client side, and then used to calculate the hash value to get the delegated credential from local repository-->\
          <!--DelegationCredIdentity>/O=KnowARC/OU=UiO/CN=squark.uio.no</DelegationCredIdentity-->\ 
          <CACertificatePath>./cacert.pem</CACertificatePath>\ 
          <CACertificatesDir>./certificates</CACertificatesDir>\ 
        </SecHandler>"); 
\end{verbatim}

@@

>>>>

%\documentclass{book}
\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
\graphicspath{{./figures/}}
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {Weizhong Qiang},
  pdftitle = {Documentation of the ARC security framework},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\tableofcontents                          %Comment if use article style
% Yet we are not using documentclass--book (article instead) here,
% we still need to use TOC.
\newpage
\renewcommand{\thefootnote}{\arabic{footnote}}

%\chapter{Technical Description} % (fold)
%\label{cha:tech_description}

\section{Introduction}
\label{sec:introduction}

@@

In the implementation of the ARC1, there is a Service Container – the Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting place for various services in application level, as well as a flexible and efficient communication mechanism.

HED contains a framework for implementing and enforcing authentication and authorization. Each Message Chain Component (MCC) or service has a common interface for implementing various authentication and authorization functionality. This functionality is implemented by using pluggable components (plug-ins) called SecHandler. The SecHandler components are C++ classes and provide method for processing messages traveling through Message Chains of the HED. Each MCC or Service  usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called ``incoming'' and ``outgoing'' respectively. It is possible for MCC or Service to implement other set of queues. Please check documentation of particular component for that particular information. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well.

\begin{figure}[ht]
\centering{{{\includegraphics[width=1.0\textwidth]{MCC_Component.pdf}}}
\caption{\label{fig:messageinMCC}\textit{There are usually two chains of SecHandlers inside the MCC or service. Each SecHandler will parse the Security Attributes which are generated by the upstream MCC/services or probably upstream SecHandlers in the same or other MCC/Service, and do message processing or authenticate or authorize the incoming/outgoing message based on the collected information. The SecHandler can also change the payload and attributes of Messsage itself. For example, the Username-Token SecHandler will insert the WSS Username Token [3] into header part of SOAP message. The PDPs are called by the SecHandlers and are supposed to make authorization decision. Here the two local PDP and one remote PDP service is just for demonstration, and one or any number of PDPs can be configured under each SecHandler.
}} }
\end{figure}

Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs. Some of implemented SecHandler components also make use of pluggable and configurable sub-modules which specifically handle various security functionalities, such as authorization, authentication, etc. The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents. Figure \ref{fig:messageinMCC} gives the structure of a MCC/Service, and the message sequence inside it.

The configuration of SecHandler components for an example ``Echo'' service is shown as below. Example Echo service is configured to use two SecHandlers, both responsible for authorization. First SecHandler uses the identity of client extracted from the incoming message to map it into local identity like local Linux username. In this case all clients are mapped to local account “test”. The second one uses two PDPs: one will compose ARC specific authorization request based on the Security Attributes collected from “incoming” message and evaluate it against the ARC specific authorization policy which is defined in local file “policy.xml”; the other will compare the X509 identity of client extracted from the incoming message against list of identities stored locally.

\begin{verbatim}
<Service name="echo" id="echo">
    <SecHandler name="identity.map" id="map" event="incoming">
        <PDP name="allow.pdp"><LocalName>test</LocalName></PDP>
    </SecHandler>
    <SecHandler name="arc.authz" id="authz" event="incoming">
        <PDP name="arc.pdp">
            <PolicyStore>
                <Location type="file">policy.xml</Location>
                <!-- other policy location-->
            </PolicyStore>
        </PDP>
        <PDP name="simplelist.pdp" location="pemittedlist.txt"/>
    </SecHandler>
</Service>
\end{verbatim}

@@

When one component (MCC or service) is loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well.

There is one simple interface (see below) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.

\begin{verbatim}
class SecHandler {
   public:
    SecHandler(Arc::Config*) {};
    virtual ~SecHandler() {};
    virtual bool Handle(Arc::Message *msg) = 0;
  };
\end{verbatim}

Class \textit{SecHandler} is an abstract class which includes a general interface called \textit{Handle} which uses Message object as argument. Any security handler implementation should inherit class \textit{SecHandler} and implement the interface according to the actual functionality. The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.

@@

Below shows the definition of abstract class PDP. The implementation could be some function which implements the interface by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result, or just by composing the policy evaluation request, invoking some remote policy decision web service and getting back the evaluation result.

\begin{verbatim}
class PDP {
   public:
    PDP(Arc::Config* cfg) { };
    virtual ~PDP() {};
    virtual bool isPermitted(Arc::Message *msg) = 0;
  };
\end{verbatim}

Class \textit{PDP} is an abstract class which includes a general interface called \textit{isPermitted} which uses Message object as argument. Any policy decision point implementation should inherit class PDP and implement the interface according to the actual functionality. The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.

@@

    \item xacml.pdp – XACML PDP
The XACML PDP will organize the security attributes into the XACML authorization request (see: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd), call the policy evaluator to evaluate the request against the policy (which is in XACML format: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd) repository, and get back the evaluation result.

@@

    \item pdpservice.invoker – PDP Service Invoker
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP or XACML PDP, except that the evaluation request and response are carried by SOAP message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation engine can be accessed remotely and maintained centrally.

@@

\begin{figure}[ht]
\centering{{{\includegraphics[width=1.0\textwidth]{Evaluator_ARC.pdf}}}
\caption{\label{fig:policyengine_ARC}The UML class diagram of the classes inside policy evaluation engine that support ARC policy} }
\end{figure}

\begin{figure}[ht]
\centering{{{\includegraphics[width=1.0\textwidth]{Evaluator_XACML.pdf}}}
\caption{\label{fig:policyengine_XACML}The UML class diagram of the classes inside policy evaluation engine that support XACML policy} }
\end{figure}

Figure \ref{fig:policyengine_ARC} and \ref{fig:policyengine_XACML} respectively show the UML class diagram about the policy evaluation engine for ARC policy and XACML policy. They show all classes and relations simultaneously for getting the overall picture.

The \textbf{Evaluator} class is the key class for policy evaluation. It accepts request evaluates it against loaded policy and returns evaluation response.

Three abstract factories - \textit{FnFactory}, \textit{AlgFactory}, \textit{AttributeFactory} - are responsible for creating the \textit{Function}, \textit{CombiningAlg} and \textit{AttributeValue} objects correspondingly. The classes inherited from \textit{CombiningAlg} class take care of implementing various combining algorithms which define relations between \textit{$<$Rule/$>$} elements in policy. The \textit{AttributeValue} type of classes are used for processing different types of \textit{$<$Attribute/$>$} and similar elements. The \textit{Function} classes take care of comparing \textit{$<$Attribute/$>$} elements of request and policy.

The \textit{Policy} class parses \textit{$<$Policy/$>$} or \textit{$<$Rule/$>$} elements and creates \textit{CombingAlg} objects according to the \textit{$<$RuleCombiningAlg/$>$} attribute of \textit{$<$Policy/$>$}, \textit{Function} objects according to the \textit{$<$Function/$>$} attribute of \textit{$<$Attribute/$>$} and \textit{AttributeValue} objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. Those objects will be used when evaluating the request.

The \textit{Request} class is responsible for parsing \textit{$<$Request/$>$} element and creates corresponding \textit{AttributeValue} objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. When evaluating, each \textit{AttributeValue} in request will be evaluated against corresponding \textit{AttributeValue} in the policy by using relevant \textit{Function}.

Due to extensible architecture of code it is relatively easy to add support for new types of \textit{AttributeValue}, \textit{Function} and \textit{CombingAlg} objects in this way supporting various types of XML based policy languages.

\subsection{Policy evaluation engine --- Support of ARC policy and request} % (fold)
\label{subsec:policy_engine_arc}

\subsubsection{Schemas for ARC policy and request} % (fold)
\label{subsubsec:arc_policy_schema}
The schema for ARC Policy is available at

@@

The hierarchy tree of ARC Policy is shown below (numbers show multiplicity of elements)

@@

\subsubsection{Basic Elements of Policy} % (fold)
\label{subsubsec:poicy_element}
There are 2 basic objects - ``policy'' and ``request''. There is 1 main actor - Evaluator. Curretly there are two types of elements in policy: \textit{Policy} and \textit{Rule}. \textit{Policy} element is made of \textit{Rule} elements.
\textit{Evaluator} matches request to policy and produces one of 4 following results:

@@

\subsubsection{Policy Matching} % (fold)
\label{subsubsec:poicy_matching}
Policy is made of Rule elements. Request is evaluated against each Rule. Each evaluation produces same results as policy evaluation described above. The results from all Rules are then combined in order to produce final result for whole policy. Results Combining Algorithm is specified in Policy. There are 26 algorithms currently:

@@

\subsubsection{Request Structure} % (fold)
\label{subsubsec:request_structure}

@@

\subsubsection{Rule Composition and Matching} % (fold)
\label{subsubsec:rule_comp_match}

@@

\subsubsection{Rule Elements Matching} % (fold)
\label{subsubsec:rule_element_match}
For elements without attributes those elements have:

@@

\subsection{Policy evaluation engine --- Support of XACML policy and request} % (fold)
\label{subsec:policy_engine_xacml}
Currently, XACML specification is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.

http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-core-spec-os.pdf

http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd

http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd

@@

For making usage of policy evaluation engine more convenient basic \textit{Evaluator} class is complemented by additional interfaces. Below are examples of steps needed to carry out policy evaluation and corresponding helper interfaces.

@@

\begin{verbatim}
    // Create object which provides an interface 
    // for loading other objects
    ArcSec::EvaluatorLoader eval_loader; 
    //Load the Evaluator 
    ArcSec::Evaluator* eval = NULL;
    // Define name of policy evaluator.
    // This one is for evaluation ARC policies
    std::string evaluator = "arc.evaluator"; 
    // If xacml evaluation engine is used, 
    // std::string evaluator = "xacml.evaluator";

    eval = eval_loader.getEvaluator(evaluator);
\end{verbatim}
b)Create the policy object:

\begin{verbatim}
    ArcSec::Policy* policy = NULL; 
    // Define type of policy – ARC policy in this case
    std::string policyclassname = "arc.policy"; 
    // If xacml policy is used, 
    // std::string policyclassname = "xacml.policy";

    // Define source from which policy to be taken
    ArcSec::SourceFile policy_source("Policy_Example.xml"); 
    // Load and parse policy
    policy = eval_loader.getPolicy(policyclassname, policy_source); 
\end{verbatim}

c)Create the request:
\begin{verbatim}
    ArcSec::Request* request = NULL;
    // Define type of request – ARC request in this case
    std::string requestclassname = "arc.request";
    // If xacml request is used, 
    // std::string requestclassname = "xacml.request";

    // Define source from which request to be taken
    ArcSec::SourceFile request_source("Request.xml");
    // Load and parse request
    request = eval_loader.getRequest(requestclassname, request_source);
\end{verbatim}

@@

The Evalutor::evaluate() method can also be feed up with both \textit{Policy}/\textit{Request} objects and their sources in any combination. See example code at http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/testinterface\_arc.cpp or http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/testinterface\_xacml.cpp for more details about usage of the interface.

@@

\section{Policy Decision Service (Charon Service)} % (fold)
\label{sec:policy_decision_service}
Policy decision service is a service implementation which configurably contains the functionality of ArcPDP or XACMLPDP. It will accept the soap request containing policy decision request and return soap response containing policy decision response.

@@

 http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.wsdl

@@

\section{Security Attributes -- how to compose policy decision request to policy evaluation engine} % (fold)
\label{sec:sec_attributes}

@@

Security Attributes represent security related information inside HED framework and store information representing various aspects needed to perform authorization decison - identity of client, requested action, targeted resource, constraint policies.

Each kind of Security Attribute is represented by own class inherited from parent SecAttr class $<$arc/message/SecAttr.h$>$.

Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method. Currently only supported format is ARC Policy/Request XML document described in Section 8.1and 8.2
Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth $<$arc/message/MessageAuth.h$>$ and MessageAuthContext $<$arc/message/Message.h$>$ storing collected attributes per request and per session correspondingly. Each attribute is assigned a name. Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins. See Section 5.2for available Collectors and corresponding Security Attributes.

Note for service developers: Services may implement own authorization algorithms. But they may use Security Atributes as well by providing instances of classes inherited from SecAttr and running them through either configured or hardcoded processors/PDPs.
Processors of Security Attributes are implemented as Policy Decision Point components. Currently there are 2 PDP components available:
Arc PDP makes use of Security Attributes containing identities of client, resource and requested action. It evaluates either all or selected set of attributes against specified Policy documents thus making it possible to enforce policies defined/selected by service providers. 
Delegation PDP is described below in Section 6.3.

@@

\subsection{Delegation Interface} % (fold)
\label{subsec:delegation_interface}

\begin{figure}
\centering{{{\includegraphics[width=0.9\textwidth]{delegation_flow.pdf}}}
\caption{\label{fig:delegation_flow}The flow diagram of delegation procedure with multiple second step and session expiration} }
\end{figure}

\begin{figure}
\centering{{{\includegraphics[width=0.9\textwidth]{delegation_flow1.pdf}}}
\caption{\label{fig:delegation_flow1}The flow diagram of delegation procedure with certificate transfered as payload of service specific message} }
\end{figure}

Delegation interface in the ARC1 is implemented using Web Service approach. Each ARC1 service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials. Here is how delegation procedure works (also shown in Figure \ref{fig:delegation_flow} and Figure \ref{fig:delegation_flow1}) :

@@

\subsection{SAMLToken SecHandler} % (fold)
\label{subsec:saml_token}

@@

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath> 
            <tls:CACertificatePath>./cacert.pem</tls:CACertificatePath> 
              <tls:CACertificatesDir>./certificates</tls:CACertificatesDir> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="soap">POST</next> 
        </Component> 
        <Component name="soap.service" id="soap"> 
            <next id="echo"/> 
        </Component> 
        <Service name="echo" id="echo"> 
            <SecHandler name='delegation.handler' id='delegation' event='incoming'> 
              <deleg:Type>x509</deleg:Type> 
              <deleg:Role>delegatee</deleg:Role> 
              <deleg:DelegationServiceEndpoint>
                https://glueball.uio.no:60000/delegation
              </deleg:DelegationServiceEndpoint> 
              <!--The credential and trusted certificates which will be used by this 
              security handler to contact with delegation service -->
              <deleg:KeyPath>./testkey-nopass.pem</deleg:KeyPath> 
              <deleg:CertificatePath>./testcert.pem</deleg:CertificatePath> 
              <deleg:CACertificatesDir>./certificates</deleg:CACertificatesDir> 
              <deleg:CACertificatePath>./testcacert.pem</deleg:CACertificatePath> 
            </SecHandler> 
            <next id="echo"/> 
            <echo:prefix>[ </echo:prefix> 
            <echo:suffix> ]</echo:suffix> 
        </Service> 
    </Chain> 
\end{verbatim}

@@

\begin{verbatim}
     Arc::XMLNode sechanlder_nd("\ 
        <SecHandler name='delegation.handler' id='delegation' event='outgoing'>\ 
          <Type>x509</Type>\ 
          <Role>delegator</Role>\ 
          <!--The endpoint of delegation service, to which this security handler will 
          delegate the credential-->\
          <DelegationServiceEndpoint>\
            https://glueball.uio.no:60005/delegation
          </DelegationServiceEndpoint>\ 
          <!--The endpoint of the peer service, to which this WS client will access-->\
          <PeerServiceEndpoint>https://selectron.uio.no:60000/echo</PeerServiceEndpoint>\ 
         <!--The key and certificate of the credential, which is be delegated to the 
         delegation service-->\
          <KeyPath>./testkey-nopass.pem</KeyPath>\ 
          <CertificatePath>./testcert.pem</CertificatePath>\ 
         <!--Another option of defining the location of credential-->\
          <!--ProxyPath>/tmp/5612d050.pem</ProxyPath-->\ 
         <!--Yet another option of defining how to get the credential; this option is 
         useful for the service which has already acquired the delegated credential 
         (this delegated credential is stored under /tmp/hash_value.pem) from the 
         delegation service, yet still needs to delegate this credential to another delegation 
         service; the DelegationCredentialIdentity can be parsed from the tls connection from 
         client side, and then used to calculate the hash value to get the delegated credential 
         from local repository-->\
          <!--DelegationCredIdentity>/O=KnowARC/OU=UiO/CN=squark.uio.no</DelegationCredIdentity-->\ 
          <CACertificatePath>./cacert.pem</CACertificatePath>\ 
          <CACertificatesDir>./certificates</CACertificatesDir>\ 
        </SecHandler>"); 
\end{verbatim}

@@

\section{SAML Attribute Authority Service} % (fold)
\label{sec:saml_aa_service}

%section saml_aa_service (end)
